---
title: "STAT 511A Homework 10"
author: "Kathleen Wendt"
date: "11/27/2019"
output:
  pdf_document: default
---

# Load packages

```{r packages, message = FALSE, warning = FALSE}
library(epitools)
library(metafor)
```

# Question 1

Bacillus Calmette-Guerin (BCG) is a vaccine for preventing tuberculosis. For this question, we will examine data from 3 studies (Vandiviere et al 1973, TPT Madras 1980, Coetzee & Berjak 1968). The data is summarized below.

A note about the BCG vaccine from Wikipedia:
The most controversial aspect of BCG is the variable efficacy found in different clinical trials that appears to depend on geography. Trials conducted in the UK have consistently shown a protective effect of 60 to 80%, but those conducted elsewhere have shown no protective effect, and efficacy appears to fall the closer one gets to the equator.

## Data tables 

### Study 1 data

```{r tb_study1_table}
tb_study1 <- matrix(c(619, 10, 2537, 8), nrow = 2, byrow = TRUE)
colnames(tb_study1) <- c("tbneg", "tbpos")
rownames(tb_study1) <- c("ctrl", "trt")
tb_study1
```

### Study 2 data

```{r tb_study2_table}
tb_study2 <- matrix(c(87892, 499, 87886, 505), nrow = 2, byrow = TRUE)
colnames(tb_study2) <- c("tbneg", "tbpos")
rownames(tb_study2) <- c("ctrl", "trt")
tb_study2
```

### Study 3 data

```{r tb_study3_table}
tb_study3 <- matrix(c(7232, 45, 7470, 29), nrow = 2, byrow = TRUE)
colnames(tb_study3) <- c("tbneg", "tbpos")
rownames(tb_study3) <- c("ctrl", "trt")
tb_study3
```

## Part 1A

Calculate the odds ratio (corresponding to TBpos for Trt vs Ctrl) for each study separately. (4 pts)

### Study 1 odds ratio

```{r study1_odds}
tb1_odds <- epitools::oddsratio(tb_study1, method = "wald")
tb1_odds$measure
```

### Study 2 odds ratio

```{r study2_odds}
tb2_odds <- epitools::oddsratio(tb_study2, method = "wald")
tb2_odds$measure
```

### Study 3 odds ratio

```{r study3_odds}
tb3_odds <- epitools::oddsratio(tb_study3, method = "wald")
tb3_odds$measure
```

## Part 1B

Use the Breslow-Day test to test for equality of odds ratios across the 3 studies. State your p-value and conclusion. Can we conclude that the odds ratios are equal across the 3 studies? Based on this test, should we combine information across studies? (4 pts)

```{r tb_array}
tb <- array(c(619, 2537, 10, 8,
              87892, 87886, 499, 505,
              7232, 7470, 45, 29),
            dim = c(2, 2, 3),
            dimnames = list(group = c("ctrl", "trt"),
                            response = c("tbneg", "tbpos"),
                            study = c("1", "2", "3")))
tb
```

```{r tb_bd_test}
cmh <- metafor::rma.mh(ai = tb[1, 1,],
                       bi = tb[1, 2,],
                       ci = tb[2, 1,],
                       di = tb[2, 2,])
cmh$BDp
```

The Breslow-Day Test for equality of odds ratios of three tuberculosis studies yielded a p-value < .001. There is evidence to suggest a difference between odds ratios by study; it is not appropriate to combine information across studies.

# Question 2

Problem 10.36 involves bomb hits during WWII. Bomb hits were recorded in n = 576 grids in a map of a region of South London.

## Create data object

```{r bomb_data}
# observations
obs <- c(229, 211, 93, 35, 8)
y <- seq(from = 0, to = 4, by = 1)
bombs <- cbind(y, obs)
bombs
```

## Part 2A

Find the sample mean ($\mu$) bomb hits per grid.

```{r bomb_mean}
# mean
muhat <- sum(obs*y)/sum(obs)
muhat
```

$\hat\mu$ = `r muhat` bomb hits per grid.

## Part 2B

Use the GOF test to test whether the number of bomb hits per grid follows the
Poisson distribution. Calculate the GOF test statistic, df, p-value and give a
conclusion using $\alpha$ = 0.05. (6 pts)

```{r bomb_prob}
# calculate the corresponding Poisson Probabilities
prob <- dpois(y, muhat)
prob
```

```{r prob_check1}
length(prob)
sum(prob)
```

```{r prob_fix}
# "fix" the final entry so that the probabilities sum to 1
prob[5] <- 1 - sum(prob[1:4])
prob
```

```{r prob_check2}
length(prob)
sum(prob)
```

```{r gof_exp}
# calculate expected values and contributions to chi-square test statistic
exp <- prob*576
x2 <- (obs-exp)^2/exp
cbind(y, obs, prob, exp, x2)
```

```{r gof_chi}
# run GOF test
gof_chi <- sum(x2)
gof_chi
```

```{r gof_df}
gof_df <- 5-2
```

```{r gof_p}
gof_p <- 1 - pchisq(gof_chi, gof_df)
gof_p
```

We fail to reject null hypothesis that data are from a Poisson distribution, _p_ = `r gof_p` > $\alpha$ = 0.05.

# Question 3

The data “PoissonData.csv” gives observations Y (counts or events) for n = 50 (units) generated from the Poisson distribution (using the rpoiss() function).

```{r poisson_data}
poisson <- readr::read_csv("PoissonData.csv")
```

## Part 3A

Calculate the sample mean and sample standard deviation. Also construct a histogram and qqplot of the data and include them in your assignment. (4 pts)

### Sample mean

```{r poisson_mu}

```

### Sample standard deviation

```{r poisson_sd}

```

### Histogram

```{r}
hist(poisson$Y)
```

### Q-Q Plot

```{r}
qqnorm(poisson$Y)
qqline(poisson$Y)
```


NOTE: Because the data comes from the Poisson distribution, you should find that the mean and the sample variance (s^2) are close. However, you should also find from the histogram and qqplot that the data looks approximately normal.

## Part 3B

Give a standard t-based 95% confidence interval for $\mu$.

## Part 3C

Following the example on CH10 Slide 106 (Death by Mule Kick CI), construct a 95%
confidence interval for µ based on the normal approximation to the Poisson
distribution. (4 pts) In order to do this, you will start by constructing a CI on the total number of events, then divide by the number of units.

NOTE: The CIs from parts B and C should be similar.
