---
title: "Pretty Example"
output: pdf_document
---

## Introduction
In this study, rats were randomly assigned to one of two groups.  For the control group n = 10 rats received a standard diet.  For the test group n = 10 rats received a calcium deficient diet.  

For both groups, a 0.15% lead-acetate solution was available to drink.  The amount of solution consumed by each rat was measured.

## Summary Statistics and Graph
```{r include = FALSE, message = FALSE}
library(tidyverse)
library(kableExtra)
ratlead <- read.csv("C:/hess/STAT511_FA11/RData/CH6_RatLead.csv")
str(ratlead)
```

Below we show the simple summary statistics and boxplots.
```{r echo = FALSE}
SumStats <- summarise(group_by(ratlead, trt), 
                      n = n(),
                      mean = mean(y), 
                      sd = sd(y),
                      SE = sd/sqrt(n))
kable(SumStats, digits = 2)
```

```{r echo = FALSE, fig.length = 4, fig.height= 3}
qplot(x = trt, y = y, geom = "boxplot", data = ratlead) +
  labs(x = "", y = "Lead consumption") +
  ggtitle("Boxplots of Lead Consumption by Group") 
```

## Statistical analysis and conclusions
R software was used for statistical analysis.  The two-sample t-test assuming equal variance was used to compare mean lead consumption for the two groups.  In other words, we test $H_0: \mu_C = \mu_D$.  

```{r include = FALSE}
out <- t.test(y ~ trt, var.equal = TRUE, data = ratlead)
```

Based on the resulting test statistic and p-value (t = `r round(out$statistic,3)`, p = `r out$p.value`), we conclude that there is a diffence between mean lead consumption for the two diets.  Rats on the deficient diet consume more of the lead solution.

