---
title: "STAT 511A Homework 11"
author: "Kathleen Wendt"
date: "12/09/2019"
output:
  pdf_document: default
---

```{r packages, include = FALSE}
library(readxl)
library(janitor)
library(tidyverse)
library(broom)
library(car)
```

```{r global_options, include = FALSE}
knitr::opts_chunk$set(fig.width = 6, 
                      fig.height = 4, 
                      fig.path = "figs/",
                      echo = FALSE, 
                      warning = FALSE, 
                      message = FALSE)
```

# Question 1

Review problem 11.22 from Ott & Longnecker regarding treadmill “time to exhaustion” (X) and 10km race times (Y).

```{r run_data}
run_data <- readxl::read_xlsx("data/ex11-22.xlsx") %>% 
  janitor::clean_names() %>% 
  dplyr::rename("race_time" = "x10_k")
```

## Part 1A: Data

Regress 10.K (Y) on Treadmill (X) and include the “summary” information in your assignment.

```{r run_fit}
run_lm <- lm(race_time ~ treadmill, data = run_data)
summary(run_lm)
```

\pagebreak

## Part 1B: Scatterplot

Create a scatterplot of 10-K vs Treadmill with fitted regression line overlaid.

```{r run_plot}
run_lm %>% 
  ggplot(aes(x = treadmill, y = race_time)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  xlab("Treadmill Time to Exhaustion") + 
  ylab("10 km Race Times (min)") + 
  ggtitle("Association between Time to Exhaustion and Race Times") +
  theme_minimal() 
```

## Part 1C: Slope

Give the estimate, 95% confidence interval and interpretation of the slope. (4 pts)

```{r tidy_run_fit}
tidy_run_lm <- broom::tidy(run_lm)
```

### Estimate

The estimate for the slope is `r tidy_run_lm$estimate[2]`.

### 95% Confidence Interval 

```{r run_slope_ci}
tidy_run_slope_ci <- broom::tidy(confint(run_lm, level = 0.95)) %>% 
  janitor::clean_names() %>% 
  dplyr::filter(rownames == "treadmill")
```

The 95% confidence interval for the slope is (`r tidy_run_slope_ci[2]`, `r tidy_run_slope_ci[3]`).

### Interpretation

A difference of a one unit increase in treadmill time to exhaustion is, on average, associated with a `r tidy_run_lm$estimate[2]` minute decrease in 10 km race time in this sample. There is a negative, linear association between treadmill time to exhaustion and 10 km race time.

## Part 1D

Give the $R^2$ value and interpretation in terms of this scenario.

$R^2$ = 0.6807

### Interpretation

68% of the variance in 10 km run times can be attributed to the model relating treadmill time to exhaustion and 10 km run times.

## Part 1E

Give the predicted 10.K time for a runner with Treadmill = 11. 

```{r run_predict}
runner <- dplyr::tibble(treadmill = 11)
runner_pred <- broom::tidy(predict(object = run_lm,
                                   runner,
                                   interval = "confidence"))
```

`r runner_pred$fit` minutes

Also provide a corresponding prediction interval.

(`r runner_pred$lwr`, `r runner_pred$upr`)

## Part 1F

Create the plots of (1) residuals vs fitted values and (2) qqplot of residuals.

```{r run_res}
run_res_fit <- dplyr::tibble(res = residuals(run_lm),
                             fit = fitted(run_lm))

run_res_fit %>%
  ggplot(aes(x = fit, y = res)) +
  geom_point() +
  xlab("Fitted Values") + 
  ylab("Residuals") +
  ggtitle("Residual Plot for Run Data") +
  theme_minimal()
```

```{r run_qq}
run_data %>% 
  ggplot(aes(sample = residuals(run_lm))) + 
  geom_qq() +
  geom_qq_line(aes(sample = residuals(run_lm)), 
               col = "red") +
  labs(title = "Normal QQ Plot for Run Data",
       x = "Theoretical Quantiles",
       y = "Sample Quantiles") +
  theme_minimal()
```

## Part 1G

Based on the plots above, subject 13 appears to be a bit of an outlier. Run a formal outlier test for this observation. Provide the p-value and make a conclusion. Note that since we identified this observation after looking at the data, a Bonferonni adjustment is appropriate.

```{r run_outlier}
run_outlier <- car::outlierTest(run_lm)
```

Using `car::outlierTest` to identify the observation with the most extreme residual, the Bonferonni p-value for observation 13 is `r run_outlier$bonf.p` > $\alpha$ = 0.05. Based on this, we fail to reject the null hypothesis that observation 13 is NOT an outlier.

\pagebreak

# Question 2

Data on age in coating Thickness (X) and Strength (Y) from an experiment involving steel are available from Canvas as Steel.csv.

## Part 2A

Regress Strength (Y) against Thick (X) and look at (1) the plot of Strength versus Thick (2) residuals versus predicted values and (3) qqplot of residuals. Include these plots in your assignment. Do the regression assumptions appear to be met? Discuss. (4 pts)

## Part 2B

Perform an F-test for “lack of fit”. Give your p-value and make a conclusion. (4 pts)

## Part 2C

Now perform a quadratic regression and create a scatterplot with the fitted curve overlaid. Include the “summary” table and plot in your assignment. This can be done with code like the following. (4 pts)

```{r}
# QFit <- lm(Strength ~ Thick + I(Thick^2), data = Steel)
# summary(QFit)
# plot(Strength ~ Thick, data = Steel)
# curve(b0 + b1*x + b2*x^2, add = TRUE)
```

Note that b0,b1,b2 need to be *replaced* with estimates from the quadratic regression.

\pagebreak

# Question 3

Review problem 11.50 from Ott & Longnecker regarding SAT Scores. 

## Part 3A

Create pairwise scatterplots for all 4 variables (Male.Verbal, Female.Verbal, Male.Math, Female.Math)

## Part 3B

Calculate pairwise (Pearson) correlations for all 4 variables. Which pair of variables has the strongest correlation? (4 pts)

## Part 3C

Provide a test of the correlation for Female.Verbal vs Female.Math. Give the p-value and conclusion in your assignment.
